Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-05 WSSSPE3/report/wg_main_user_community.tex
Update wg_main_user_community.tex
added discussion section

Several books have been written about software communities:
\begin{itemize}
\item "Art of Community" by Jono Bacon. We could consider distilling this for scientific software.

\subsubsection{Discussion}
\todo{short-ish text here}

>>>>

Several books have been written about software communities:
\begin{itemize}
\item "Art of Community" by Jono Bacon. We could consider distilling this for scientific software.
\item Iain Larmour, from EPSRC inthe UK [[https://www.epsrc.ac.uk/][EPSRC]] (not sure who from mentioned UK Collaborative Computational Projects ([[http://ccp.ac.uk][CCP]])
\end{itemize}

\subsubsection{Discussion}
%\todo{short-ish text here}

Discussion revolved around a few questions: what is the benefit of having a "community" for software sustainability, what practices and circumstances lead to having a community, how can funding help or hinder this process, and perhaps most importantly how can best practices be described and distilled into a document that can help new projects.

The benefits of having a community that were brought up were considered largely obvious. In addition to having advocates for the software, and a possible source of ``free'' contributions to the codebase, the community becomes a good source for requirements, feedback, and metrics. The software community can also act as "cheerleaders" who convince funders or other potential users to fund/use the software, and thus help sustain the software.

Practices and circumstances that lead to a community are first, that the software offers value. But in addition to this, a community will be much more likely to form if they receive (expert) support when they have questions. Additional contributing factors are good usability (not always needed), and an open development process such as IPython developer meetings on YouTube. It was also pointed out that an evangelist for the project, not necessarily but often one of the developers, can often make a big difference.

Funding can help the process by encouraging both value to the community and high-quality user support. Only providing funding for the software development may create good software, but with less likelihood to have a real community. It was discussed that federal laboratories are a good incubator for software communities, and that a general facility like EarthCube is too dispersed to really make a community. Also, domain-specific groups within laboratories or universities might provide as an incubator for software communitites.

In describing best practices, the group discussed the different modes for starting a scientific software project: building on an existing product that needs improving, recognizing an unsatisfied need of an existing community, or creating a new solution to a need not yet recognized by the community. The group also thought that the existing books on software communities would need to be evaluated in light of differences between Science Software projets and general OSS projects in terms of scale, science, acknowledgement and credit, and funding models.
